---
title: "writeup"
author: "Kaleb Cervantes"
date: "September 19, 2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The following code chunk loads the neccessary libraries:
```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(lubridate)
```

The next code chunk loads the game portion of the kaggle dataset.
```{r}
# loads game dataset with as many variables in ideal types as possible
Twitch_game_data <- read_csv(
  "data/Twitch_game_data.csv",
  col_types = cols(
    Rank = col_integer(),
    Month = col_integer(),
    Year = col_integer(),
    Hours_watched = col_integer(), 
    Hours_Streamed = col_character(), 
    Peak_viewers = col_integer(),
    Peak_channels = col_integer(),
    Streamers = col_integer(),
    Avg_viewers = col_integer(),
    Avg_channels = col_integer()
  )
)

# converts Month variable to a date type
Twitch_game_data$Month <- (
  Twitch_game_data %>%
  select(Year, Month) %>%
  transmute(tmp = make_date(year=Year, month=Month))
)$tmp

# remove Year variable
Twitch_game_data <- within(Twitch_game_data, rm(Year))

# changes Hours_Streamed from char to integer
Twitch_game_data$Hours_Streamed <- sapply(
  
  #remove non numeric characters
  sub(" hours", "", Twitch_game_data$Hours_Streamed),
  
  as.integer
)
```

We can also load in a data set that contains some global data to get a better idea of how relatively popular these categories are.

```{r}
# loads global dataset with as many variables in ideal types as possible
Twitch_global_data <- read_csv(
  "data/Twitch_global_data.csv",
  col_types = cols(
    year = col_integer(),
    Month = col_integer(),
    Hours_watched = col_number(),
    Avg_viewers = col_integer(),
    Peak_viewers = col_integer(),
    Streams = col_integer(),
    Avg_channels = col_integer()
  )
)

# converts Month variable to a date type
Twitch_global_data$Month <- (
  Twitch_global_data %>%
    select(year, Month) %>%
    transmute(tmp = make_date(year=year, month=Month))
)$tmp

# remove Year variable
Twitch_global_data <- within(Twitch_global_data, rm(year))

# converts string column to integer
Twitch_global_data$Games_streamed <- sapply(
  
  # removes comma from string
  sub(",", "", Twitch_global_data$Games_streamed),
  
  as.integer
)
```

In order to get an idea about the relative popularity of each game, I decided to add a column that gives the ratio of hours watched views for the game over the site.

```{r}
Twitch_game_data$ratio_hours_watched <- (
  
  # subset game data
  Twitch_game_data %>%
    select(Rank, Month, Hours_watched) %>%
    merge(
      
      # subset global data
      Twitch_global_data %>%
        select(Month, Hours_watched),
      
      # inner join by month
      by = "Month",
      
      # rename variables to avoid ambiguity
      suffixes = c("_game", "_global")
    ) %>%
    
    # create temporary column
    transmute(tmp = Hours_watched_game / Hours_watched_global)
  
  # extract temporary column
)$tmp
```

I was asked by someone in a Discord server about what games were in the top 5 popularity for Twitch. To do this, I decided to subset the game data and visualize with a barplot.
```{r}
# filters out top 5 for each month
top5 <- subset(Twitch_game_data, Rank <= 5)

# bar plot
ggplot(top5) +
  
  # factor is needed or else color will be grey
  geom_bar(aes(x=Game, fill=factor(Rank))) +
  labs(fill="Monthly Popularity") +
  
  # more visible on laptop
  coord_flip()
```